---
title: "A1"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(ggplot2)
setwd("C:\\Users\\RAJAAB\\Downloads")
activityData <- read.csv('activity.csv')

```

#What is mean total number of steps taken per day?
```{r}
stepsByDay <- tapply(activityData$steps, activityData$date, sum, na.rm=TRUE)
stepsByDay
```
#Make a histogram of the total number of steps taken each day
```{r}
qplot(stepsByDay, xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)

```
#2. Calculate and report the mean and median total number of steps taken per day
```{r}
stepsByDayMean <- mean(stepsByDay)
print(stepsByDayMean)
stepsByDayMedian <- median(stepsByDay)
print(stepsByDayMedian)
```

#What is the average daily activity pattern?

```{r}
averageStepsPerTimeBlock <- aggregate(x=list(meanSteps=activityData$steps), by=list(interval=activityData$interval), FUN=mean, na.rm=TRUE)

```
#1. Make a time series plot
```{r pressure, echo=FALSE}
ggplot(data=averageStepsPerTimeBlock, aes(x=interval, y=meanSteps)) +
  geom_line() +
  xlab("5-minute interval") +
  ylab("average number of steps taken") 

```
#2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}

mostSteps <- which.max(averageStepsPerTimeBlock$meanSteps)
mostSteps

```
```{r}
timeMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsPerTimeBlock[mostSteps,'interval'])
timeMostSteps
```

#Imputing missing values
#1. Calculate and report the total number of missing values in the dataset
```{r}

numMissingValues <- length(which(is.na(activityData$steps)))
numMissingValues
```
#2. Devise a strategy for filling in all of the missing values in the dataset.
#3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}

library(Hmisc)
activityDataImputed <- activityData
activityDataImputed$steps <- impute(activityData$steps, fun=mean)
```

#4. Make a histogram of the total number of steps taken each day
```{r}

stepsByDayImputed <- tapply(activityDataImputed$steps, activityDataImputed$date, sum)
qplot(stepsByDayImputed, xlab='Total steps per day (Imputed)', ylab='Frequency using binwith 500', binwidth=500)

```

#... and Calculate and report the mean and median total number of steps taken per day.
```{r}

stepsByDayMeanImputed <- mean(stepsByDayImputed)
stepsByDayMedianImputed <- median(stepsByDayImputed)
```
#Are there differences in activity patterns between weekdays and weekends?
#1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
activityDataImputed$dateType <-  ifelse(as.POSIXlt(activityDataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```


#2. Make a panel plot containing a time series plot
```{r}

averagedActivityDataImputed <- aggregate(steps ~ interval + dateType, data=activityDataImputed, mean)
ggplot(averagedActivityDataImputed, aes(interval, steps)) + 
  geom_line() + 
  facet_grid(dateType ~ .) +
  xlab("5-minute interval") + 
  ylab("avarage number of steps")


```


